---
title: "Problem_set_1_MP"
author: "Manon Pillot"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align ="center"
)
```

```{r, libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(vroom)
library(skimr)
library(sf)
library(extrafont)
library(gapminder)
library(ggrepel)
library(patchwork)
library(remotes)
library(nycdogs)
library(ggplot2)
library(ggmap)
library(gridExtra)
library(pander)
library(here)
library(leaflet)
library(tmap)
library(tmaptools)
library(hrbrthemes)
library(mapview)
library(viridis)
library(plotly)
library(htmlwidgets)
library(scales)
library(svglite)
library(treemapify)
library(ggalluvial)
library(ggridges)
library(ggthemes)
library(packcircles)
library(ggiraph)
library(ggtext)
library(gghighlight)
library(knitr)
library(RColorBrewer)
library(stringr)
library(zoo)
library(httr)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(rgeos)
loadfonts(device="win")
#library(MASS)
# font_import()
```

# import data

```{r, import data}
# import world death causes data 
data_1 <- readr::read_csv('data/annual-number-of-deaths-by-cause.csv')
# import 15-49 age group world death causes data
data_2 <- readr::read_csv('data/causes-of-death-in-15-49-year-olds.csv')
# import world countries population data
data_pop <- read_csv("data/pop.csv")
# import world countries homicide cases amount data
data_hm_n <- read_csv(file = "data/homicide_count.csv")
# import world homicide rate data 
homicide_raw <- vroom("data/homicide-rate-1990-vs-2019.csv")
# import world continent homicide rate data
homicide_raw_long <- vroom("data/data_countries.csv") 
# import world bank and economic data 
world_bank <- vroom("data/world_bank_data.csv")
# import world happiness data & download data to temporary file
url <- "https://happiness-report.s3.amazonaws.com/2021/DataPanelWHR2021C2.xls"
httr::GET(url, write_disk(happiness.temp <- tempfile(fileext = ".xls")))
# import data with information for victim and the affected age group 
victim_age_data <- read.csv("data/data_countries_age.csv")
# import data for intentional homicide
intentional_homicide_data_all <- read.csv("data/data_cts_intentional_homicide.csv")
```

# Human unnatural death rank

## Homicide is the 17th death cause in the world (2019 Data)

```{r, world_death_causes_analysis_bar_chart1}
# Select data needed
graph_1 <- data_1 %>% 
           filter(Year == "2019") %>%
           filter(Entity == "World") %>% 
           rename(Meningitis = `Deaths - Meningitis - Sex: Both - Age: All Ages (Number)`,
            Dementias = `Deaths - Alzheimer's disease and other dementias - Sex: Both - Age: All Ages (Number)`,
            "Cardiovascular diseases" = `Deaths - Cardiovascular diseases - Sex: Both - Age: All Ages (Number)`,
            "Parkinson's disease" = `Deaths - Parkinson's disease - Sex: Both - Age: All Ages (Number)`,
            "Nutritional deficiencies" = `Deaths - Nutritional deficiencies - Sex: Both - Age: All Ages (Number)`,
            Malaria = `Deaths - Malaria - Sex: Both - Age: All Ages (Number)`,
            Drowning = `Deaths - Drowning - Sex: Both - Age: All Ages (Number)`,
            Homicide = `Deaths - Interpersonal violence - Sex: Both - Age: All Ages (Number)`,
            "Maternal disorders" = `Deaths - Maternal disorders - Sex: Both - Age: All Ages (Number)`,
            "HIV/AIDS" = `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Number)`,
            "Drug use disorders" = `Deaths - Drug use disorders - Sex: Both - Age: All Ages (Number)`,
            Tuberculosis = `Deaths - Tuberculosis - Sex: Both - Age: All Ages (Number)`,
            "Lower respiratory infections" = `Deaths - Lower respiratory infections - Sex: Both - Age: All Ages (Number)`,
            "Neonatal disorders" = `Deaths - Neonatal disorders - Sex: Both - Age: All Ages (Number)`,
            "Alcohol use disorders" = `Deaths - Alcohol use disorders - Sex: Both - Age: All Ages (Number)`,
            "Kidney disease" = `Deaths - Chronic kidney disease - Sex: Both - Age: All Ages (Number)`,
            Suicide = `Deaths - Self-harm - Sex: Both - Age: All Ages (Number)`,
            "Natural disasters" = `Deaths - Exposure to forces of nature - Sex: Both - Age: All Ages (Number)`,
            "Diarrheal diseases" = `Deaths - Diarrheal diseases - Sex: Both - Age: All Ages (Number)`,
            Heat = `Deaths - Environmental heat and cold exposure - Sex: Both - Age: All Ages (Number)`,
            Cancers = `Deaths - Neoplasms - Sex: Both - Age: All Ages (Number)`,
            Conflict = `Deaths - Conflict and terrorism - Sex: Both - Age: All Ages (Number)`,
            Diabetes = `Deaths - Diabetes mellitus - Sex: Both - Age: All Ages (Number)`,
            Poisonings = `Deaths - Poisonings - Sex: Both - Age: All Ages (Number)`,
            Hepatitis = `Deaths - Acute hepatitis - Sex: Both - Age: All Ages (Number)`,
            "Protein-energy malnutrition" = `Deaths - Protein-energy malnutrition - Sex: Both - Age: All Ages (Number)`,
            "Road injuries" = `Deaths - Road injuries - Sex: Both - Age: All Ages (Number)`,
            "Respiratory diseases" = `Deaths - Chronic respiratory diseases - Sex: Both - Age: All Ages (Number)`,
            "Liver diseases" = `Deaths - Cirrhosis and other chronic liver diseases - Sex: Both - Age: All Ages (Number)`,
            "Digestive diseases" = `Deaths - Digestive diseases - Sex: Both - Age: All Ages (Number)`,
            Fire = `Deaths - Fire, heat, and hot substances - Sex: Both - Age: All Ages (Number)`) %>% 
            select(Meningitis, "Maternal disorders", Hepatitis, Suicide, Homicide, Drowning, Malaria, Fire, Cancers, "Digestive diseases", "Liver diseases", "Alcohol use disorders", "Respiratory diseases", "Kidney disease", "Cardiovascular diseases", "Drug use disorders", "Nutritional deficiencies", "Lower respiratory infections", Diabetes, "Protein-energy malnutrition", Heat, "Diarrheal diseases", "Road injuries", Tuberculosis, "HIV/AIDS", "Parkinson's disease", "Neonatal disorders", Conflict, Poisonings, Dementias, "Natural disasters") %>% 
            pivot_longer(everything()) %>% 
            rename(Cause_of_Death = name,
                   Total = value) %>% 
            mutate(Homicide = ifelse(Cause_of_Death == "Homicide", TRUE, FALSE)) %>% 
            slice_max(Total, n = 20)

#define colours to use: grey for everything, tomato for controlled drugs
my_colours <- c("grey80", "#CB102D")

#create plot
graph_1_svg <- ggplot(graph_1, aes(x=fct_reorder(Cause_of_Death, Total, max),
                y = Total,
                fill = Homicide))+
       scale_fill_manual(values = my_colours)+
       geom_bar(stat = "identity", show.legend = FALSE, width = 0.8)+
       theme_minimal()+
       scale_y_continuous(labels=comma)+
       coord_flip()+
  
#labels
  labs(x = '', 
       y = 'Number of death')+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_text(color = "black", size = 18),
        axis.text.x=element_text(color = "black", size = 18),
        text=element_text(family="Roboto"),
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        )

graph_1_svg

ggsave(file="graph_1.svg", plot=graph_1_svg, width=9, height=9)
```

# Human unnatural death rank in 15-49 age group

## Homicide is the N°9th death cause for people between 15 and 49 (2019 Data)

```{r, world_death_causes_analysis_bar_chart2}
# Select data needed
graph_2 <- data_2 %>% 
           filter(Year == "2019") %>%
           filter(Entity == "World") %>%
          rename(Hepatitis = `Deaths - Acute hepatitis - Sex: Both - Age: 15-49 years (Number)`,
            Suicide = `Deaths - Self-harm - Sex: Both - Age: 15-49 years (Number)`,
            Homicide = `Deaths - Interpersonal violence - Sex: Both - Age: 15-49 years (Number)`,
            Drowning = `Deaths - Drowning - Sex: Both - Age: 15-49 years (Number)`,
            Malaria = `Deaths - Malaria - Sex: Both - Age: 15-49 years (Number)`,
            Fire = `Deaths - Fire, heat, and hot substances - Sex: Both - Age: 15-49 years (Number)`,
            Cancers = `Deaths - Neoplasms - Sex: Both - Age: 15-49 years (Number)`,
            "Digestive diseases" = `Deaths - Digestive diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Alcohol use disorders" = `Deaths - Alcohol use disorders - Sex: Both - Age: 15-49 years (Number)`,
            "Liver diseases" = `Deaths - Cirrhosis and other chronic liver diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Respiratory diseases" = `Deaths - Chronic respiratory diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Kidney disease" = `Deaths - Chronic kidney disease - Sex: Both - Age: 15-49 years (Number)`,
            "Cardiovascular diseases" = `Deaths - Cardiovascular diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Drug use disorders" = `Deaths - Drug use disorders - Sex: Both - Age: 15-49 years (Number)`,
            "Nutritional deficiencies" = `Deaths - Nutritional deficiencies - Sex: Both - Age: 15-49 years (Number)`,
            "Lower respiratory infections" = `Deaths - Lower respiratory infections - Sex: Both - Age: 15-49 years (Number)`,
            Diabetes = `Deaths - Diabetes mellitus - Sex: Both - Age: 15-49 years (Number)`,
            "Protein-energy malnutrition" = `Deaths - Protein-energy malnutrition - Sex: Both - Age: 15-49 years (Number)`,
            "Natural disasters" = `Deaths - Exposure to forces of nature - Sex: Both - Age: 15-49 years (Number)`,
            Heat = `Deaths - Environmental heat and cold exposure - Sex: Both - Age: 15-49 years (Number)`,
            "Diarrheal diseases" = `Deaths - Diarrheal diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Road injuries" = `Deaths - Road injuries - Sex: Both - Age: 15-49 years (Number)`,
            Tuberculosis = `Deaths - Tuberculosis - Sex: Both - Age: 15-49 years (Number)`,
            "HIV/AIDS" = `Deaths - HIV/AIDS - Sex: Both - Age: 15-49 years (Number)`,
            Dementias = `Deaths - Alzheimer's disease and other dementias - Sex: Both - Age: 15-49 years (Number)`,
            "Parkinson's disease" = `Deaths - Parkinson's disease - Sex: Both - Age: 15-49 years (Number)`) %>%
            select(Hepatitis, Suicide, Homicide, Drowning, Malaria, Fire, Cancers, "Digestive diseases", "Liver diseases", "Alcohol use disorders", "Respiratory diseases", "Kidney disease", "Cardiovascular diseases", "Drug use disorders", "Nutritional deficiencies", "Lower respiratory infections", Diabetes, "Protein-energy malnutrition", Heat, "Diarrheal diseases", "Road injuries", Tuberculosis, "HIV/AIDS", "Parkinson's disease", "Natural disasters") %>%
            pivot_longer(everything()) %>% 
            rename(Cause_of_Death = name,
                   Total = value) %>% 
            mutate(Homicide = ifelse(Cause_of_Death == "Homicide", TRUE, FALSE)) %>% 
            slice_max(Total, n = 20)

#define colours to use: grey for everything, tomato for controlled drugs
my_colours <- c("grey80", "#CB102D")

#create plot
graph_2_svg <- ggplot(graph_2, aes(x=fct_reorder(Cause_of_Death, Total, max),
                y = Total,
                fill = Homicide))+
       scale_fill_manual(values = my_colours)+
       geom_bar(stat = "identity", show.legend = FALSE, width = 0.8)+
       theme_minimal()+
       scale_y_continuous(labels=comma)+
       coord_flip()+
  
#labels
  labs(x = '', 
       y = 'Number of death')+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_text(color = "black", size = 18),
        axis.text.x=element_text(color = "black", size = 18),
        text=element_text(family="Roboto"),
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        )


graph_2_svg

ggsave(file="graph_2.svg", plot=graph_2_svg, width=9, height=9)
```

# Homicide amount jumps to the first place in Columbia

## Homicide is primary cause of death for people between 15 and 49 in Colombia(2019 Data)

```{r, world_death_causes_analysis_bar_chart3}
# Select data needed
graph_3 <- data_2 %>% 
           filter(Year == "2019") %>%
           filter(Entity == "Colombia") %>% 
          rename(Hepatitis = `Deaths - Acute hepatitis - Sex: Both - Age: 15-49 years (Number)`,
            Suicide = `Deaths - Self-harm - Sex: Both - Age: 15-49 years (Number)`,
            Homicide = `Deaths - Interpersonal violence - Sex: Both - Age: 15-49 years (Number)`,
            Drowning = `Deaths - Drowning - Sex: Both - Age: 15-49 years (Number)`,
            Malaria = `Deaths - Malaria - Sex: Both - Age: 15-49 years (Number)`,
            Fire = `Deaths - Fire, heat, and hot substances - Sex: Both - Age: 15-49 years (Number)`,
            Cancers = `Deaths - Neoplasms - Sex: Both - Age: 15-49 years (Number)`,
            "Digestive diseases" = `Deaths - Digestive diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Alcohol use disorders" = `Deaths - Alcohol use disorders - Sex: Both - Age: 15-49 years (Number)`,
            "Liver diseases" = `Deaths - Cirrhosis and other chronic liver diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Respiratory diseases" = `Deaths - Chronic respiratory diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Kidney disease" = `Deaths - Chronic kidney disease - Sex: Both - Age: 15-49 years (Number)`,
            "Cardiovascular diseases" = `Deaths - Cardiovascular diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Drug use disorders" = `Deaths - Drug use disorders - Sex: Both - Age: 15-49 years (Number)`,
            "Nutritional deficiencies" = `Deaths - Nutritional deficiencies - Sex: Both - Age: 15-49 years (Number)`,
            "Lower respiratory infections" = `Deaths - Lower respiratory infections - Sex: Both - Age: 15-49 years (Number)`,
            Diabetes = `Deaths - Diabetes mellitus - Sex: Both - Age: 15-49 years (Number)`,
            "Protein-energy malnutrition" = `Deaths - Protein-energy malnutrition - Sex: Both - Age: 15-49 years (Number)`,
            "Natural disasters" = `Deaths - Exposure to forces of nature - Sex: Both - Age: 15-49 years (Number)`,
            Heat = `Deaths - Environmental heat and cold exposure - Sex: Both - Age: 15-49 years (Number)`,
            "Diarrheal diseases" = `Deaths - Diarrheal diseases - Sex: Both - Age: 15-49 years (Number)`,
            "Road injuries" = `Deaths - Road injuries - Sex: Both - Age: 15-49 years (Number)`,
            Tuberculosis = `Deaths - Tuberculosis - Sex: Both - Age: 15-49 years (Number)`,
            "HIV/AIDS" = `Deaths - HIV/AIDS - Sex: Both - Age: 15-49 years (Number)`,
            Dementias = `Deaths - Alzheimer's disease and other dementias - Sex: Both - Age: 15-49 years (Number)`,
            "Parkinson's disease" = `Deaths - Parkinson's disease - Sex: Both - Age: 15-49 years (Number)`) %>% 
            select(Hepatitis, Suicide, Homicide, Drowning, Malaria, Fire, Cancers, "Digestive diseases", "Liver diseases", "Alcohol use disorders", "Respiratory diseases", "Kidney disease", "Cardiovascular diseases", "Drug use disorders", "Nutritional deficiencies", "Lower respiratory infections", Diabetes, "Protein-energy malnutrition", "Natural disasters", Heat, "Diarrheal diseases", "Road injuries", "Tuberculosis", "HIV/AIDS", Dementias, "Parkinson's disease") %>% 
            pivot_longer(everything()) %>% 
            rename(Cause_of_Death = name,
                   Total = value) %>% 
            mutate(Homicide = ifelse(Cause_of_Death == "Homicide", TRUE, FALSE)) %>% 
            slice_max(Total, n = 20)

#define colours to use: grey for everything, tomato for controlled drugs
my_colours <- c("grey80", "#CB102D")

#create plot
graph_3_svg <- ggplot(graph_3, aes(x=fct_reorder(Cause_of_Death, Total, max),
                y = Total,
                fill = Homicide))+
       scale_fill_manual(values = my_colours)+
       geom_bar(stat = "identity", show.legend = FALSE, width = 0.8)+
       theme_minimal()+
       scale_y_continuous(labels=comma)+
       coord_flip()+
  
#labels
  labs(x = '', 
       y = 'Number of death')+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_text(color = "black", size = 18),
        axis.text.x=element_text(color = "black", size = 18),
        text=element_text(family="Roboto"),
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        )

graph_3_svg

ggsave(file="graph_3.svg", plot=graph_3_svg, width=9, height=9)
```

# Total word countries homicide case amount

## Brasil has the most amount of homicide(2019 Data)

```{r, world_homicide_amount_analysis_map1}
# clean map and population data
map <- ne_countries(scale = "medium", returnclass = "sf") %>%
  dplyr::select(name, iso_a3, geometry) %>%
  filter(!name %in% c("Antarctica"))
data_pop<- data_pop %>% 
  clean_names() %>% 
  rename(country=country_name, pop=x2019, code =country_code)

# left join map and pop data
map_pop<-map %>% 
  rename(country=name, code=iso_a3) %>% 
  left_join(data_pop %>% 
             select(pop,country), by ="country") %>% 
  left_join(data_pop %>% select(pop,code), by = "code")%>% 
  mutate(popu= if_else(is.na(pop.x),0,pop.x)+if_else(is.na(pop.y),0,pop.y) -
           if_else((!is.na(pop.x)&!is.na(pop.y)),pop.x,0)) %>% 
  mutate(pop = if_else(popu==0, as.numeric(NA), popu)) %>% 
  select(-pop.x, -pop.y,-popu)
           
# clean word countries homicide case amount data
data_n <-data_hm_n %>% 
  clean_names() %>% 
  filter(year=="2019") %>% 
  rename(country=d)

# left join it with population data 
df_count1 <- map_pop %>% 
  left_join(
    data_n %>% 
      select(country,value), by =  "country") 
df_count2<-df_count1 %>% 
  filter(is.na(code)==FALSE) %>% 
  left_join(
  data_n %>% select(code, value), by =  "code") %>% 
  rename(value=value.y) %>% 
  st_drop_geometry()
df_count<- df_count1 %>% 
  left_join(df_count2 %>% select (country,value ), by ="country") %>% 
  mutate(value1= if_else(is.na(value.x),0,value.x+0.001)+if_else(is.na(value.y),0,value.y+0.001)   - if_else((!is.na(value.x)&!is.na(value.y)),value.x+0.001,0)) %>% 
  mutate(value = if_else(value1==0, as.numeric(NA), round(value1))) %>% 
  select(-value1, -value.x, -value.y)

# mutate homicide rate data 
df_rate <- df_count %>% 
  mutate(value_rate= 100000*value/pop)

# draw map for amount
base_map <- ggplot(data = df_count) +
  geom_sf(
    mapping = aes(
      geometry = geometry, #use Natural Earth World boundaries
      fill = value
    ),
    colour = "#4c4c4c",      # white borders between regions
  )+
  scale_fill_gradient(
    low = "#ffffff",
    high = '#c40014',
    na.value = "white")+
  labs(
    title = "", fill=""
  )+
  theme_void()+
  coord_sf(crs = "+proj=robin")+
  theme(legend.position="bottom",
          legend.direction = "horizontal", legend.key.size = unit(0.5, "cm"),
  legend.key.width = unit(2,"cm") )+
  NULL
base_map
```

# Total word countries homicide rate(peer 100000 people) in 2019

## Venezuela owns higest homicide rate (2019 Data)

```{r, world_homicide_rate_analysis_map2}
# draw map for rate 
base_map_rate <- ggplot(data = df_rate) +
  geom_sf(
    mapping = aes(
      geometry = geometry, #use Natural Earth World boundaries
      fill = value_rate 
    ),
     colour = "#4c4c4c",      # white borders between regions
  )+
  scale_fill_gradient(
    low = "#fffbf9",
    high = '#c40014',
    na.value = "white")+
  labs(
    title = "", fill=""
  )+
  theme_void()+
  coord_sf(crs = "+proj=robin")+
  theme(legend.position="bottom",
          legend.direction = "horizontal", legend.key.size = unit(0.5, "cm"),
  legend.key.width = unit(2,"cm") )+
  NULL

base_map_rate
```

# 1990 vs 2019 global experience countries homicide rates compare?

## conclusion (1990 and 2019 Data)

```{r, countries_homicide_rate_analysis_scatter_plots}
# clean data
homicide_raw<- homicide_raw %>%  clean_names()
glimpse(homicide_raw)

# set filters
years <- c(1990, 2019)
countries <- c("GBR", "USA", "ZAF", "COL", "TUR", "SWE", "FRA")
year_names <- c("year_1990", "year_2019", "year_1990", "year_2019", "year_1990", "year_2019", "year_1990", "year_2019",
                "year_1990", "year_2019", "year_1990", "year_2019", "year_1990", "year_2019")

# clean data
homicide_years <- homicide_raw %>% 
  rename(year = year_3) %>% 
  rename(homicide_rate = deaths_interpersonal_violence_sex_both_age_age_standardized_rate_4) %>% 
  rename(country = entity) %>% 
  select(country, code, year, homicide_rate) %>% 
  filter(code %in% countries) %>% 
  filter(year %in% years) %>% 
  mutate(year = year_names)

# transpose data
homicide_years <- homicide_years %>% 
  pivot_wider(names_from = year, values_from = homicide_rate)

# plot figure
years_plot <- homicide_years %>% 
  ggplot(aes(x = year_1990, y = year_2019, label = country)) +
  geom_point(size = 3.5, alpha = 1, shape = 21, fill = "#CB102D", color = "black")+
  geom_segment(aes(x = 0, y = 0, xend = 80, yend = 80), 
               linetype = 2, size = 0.5, color = "black")+
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 80), size = 0.5, color = "black")+
  geom_segment(aes(x = 0, y = 0, xend = 80, yend = 0), size = 0.5, color = "black")+
  #geom_label_repel()+
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text=element_text(color = "black", size = 18),
        text=element_text(family="Roboto"),
        plot.caption = element_text(color = "grey", hjust = 0, vjust = -5),
        )

years_plot

# zoom on european countries
years_plot_zoom <- homicide_years %>% 
  ggplot(aes(x = year_1990, y = year_2019, label = country)) +
  geom_point(size = 5, alpha = 1, shape = 21, fill = "#CB102D", color = "black")+
  geom_segment(aes(x = 0, y = 0, xend = 2, yend = 2), 
               linetype = 2, size = 0.5, color = "black")+
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 2), size = 0.5, color = "black")+
  geom_segment(aes(x = 0, y = 0, xend = 2, yend = 0), size = 0.5, color = "black")+
  #geom_label_repel()+
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text=element_text(color = "black", size = 30),
        text=element_text(family="Roboto"),
        plot.caption = element_text(color = "grey", hjust = 0, vjust = -5),
        )+
  ylim(0,2)+
  xlim(0,2)

years_plot_zoom


ggsave(file="years.svg", plot=years_plot, width=7, height=7)
ggsave(file="years_zoom.svg", plot=years_plot_zoom, width=7, height=7) 
```

# Crosses global experience continents homicide rates and cases amount compare

## Conclusion ? (2019 Data)

```{r,continents_homicide_rate_analysis_bar_charts}
# clean data
homicide_raw_long <-homicide_raw_long%>% 
  clean_names()

#select relevant continents
continents <- c("Europe", "Asia", "Americas", "Oceania", "Africa")

#filter all the data
homicide_continents <- homicide_raw_long %>% 
  filter(year == 2019) %>% 
  filter(dimension == "Total") %>% 
  filter(indicator == "Victims of intentional homicide") %>% 
  filter(sex == "Total") %>% 
  filter(age == "Total") %>% 
  filter(region %in% continents) %>% 
  filter(unit_of_measurement == "Counts") %>% 
  group_by(region) %>% 
  summarize(homicide_count = sum(value)) %>% 
  mutate(region = fct_reorder(region, homicide_count, .desc = TRUE))

#plot counts
counts_continent_plot <- homicide_continents %>% 
  ggplot(aes(x = region, y = homicide_count))+
  geom_text(aes(x=region, y=homicide_count, label = homicide_count), size = 6, vjust = -1)+
  geom_col(fill = "#CB102D", width = 0.08)+
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.text.x=element_text(color = "black", size = 18),
        text=element_text(family="Roboto"),
        panel.grid = element_blank(),
        plot.caption = element_text(color = "grey", hjust = 0, vjust = -5),
        )+
  ylim(0,150000)

counts_continent_plot

# add population per continent and calculate rate
homicide_continents_prop <- homicide_continents %>% 
  mutate(population = c(1340598147, 1022831978, 4641054775, 747636026, 43111704)) %>% 
  mutate(homicide_rate = homicide_count/population * 1000000) %>% 
  mutate(region = fct_reorder(region, homicide_count, .desc = TRUE))
  

# plot it

rate_continent_plot <- homicide_continents_prop %>% 
  ggplot(aes(x = region, y = homicide_rate))+
  geom_col(fill = "#CB102D", width = 0.08)+
  geom_text(aes(x=region, y=homicide_rate, label = round(homicide_rate,0)), size = 6, vjust = -1)+
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.text.x=element_text(color = "black", size = 18),
        text=element_text(family="Roboto"),
        panel.grid = element_blank(),
        plot.caption = element_text(color = "grey", hjust = 0, vjust = -5),
        )+
  ylim(0,145)

rate_continent_plot

# Save files
ggsave(file="count.svg", plot=counts_continent_plot, width=8, height=4)
ggsave(file="rate.svg", plot=rate_continent_plot, width=8, height=4)


# Boxplots for distribution

homicide_continents_box <- homicide_raw_long %>% 
  filter(year == 2019) %>% 
  filter(dimension == "Total") %>% 
  filter(indicator == "Victims of intentional homicide") %>% 
  filter(sex == "Total") %>% 
  filter(age == "Total") %>% 
  filter(region %in% continents) %>% 
  filter(unit_of_measurement == "Rate per 100,000 population") 

box_rate <- homicide_continents_box %>% 
  ggplot(aes(x = reorder(region, -value, FUN = median), y = value)) +
  geom_boxplot(width = 0.5)+
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x=element_text(color = "black", size = 18),
        axis.text.y=element_text(color = "black", size = 15),
        text=element_text(family="Roboto"),
        panel.grid.major.x = element_blank(),
        )

box_rate

ggsave(file="box.svg", plot=box_rate, width=9, height=3.7)

  #group_by(region) %>% 
  #summarize(homicide_count = sum(value)) %>% 
  #mutate(region = fct_reorder(region, homicide_count, .desc = TRUE))


```

# Calculaton for Global experience countries comparison in gender 

## Conclusion (2019 Data)

```{r, countries_homicide_rate_analysis_in_gender_table}
# Calculate summary statistics for graph
genders <- c("Female", "Male")

# calculate count by gender
gender_all <- homicide_raw_long %>% 
  filter(year == 2019) %>% 
  filter(dimension == "Total") %>% 
  filter(indicator == "Victims of intentional homicide") %>% 
  filter(sex %in% genders) %>% 
  filter(age == "Total") %>% 
  filter(region %in% continents) %>% 
  filter(unit_of_measurement == "Counts") %>% 
  group_by(sex) %>% 
  summarize(homicide_count = sum(value))

# print proportion
percentage_all <- gender_all[1,2]/(gender_all[1,2]+gender_all[2,2]) 

percentage_all %>% 
  rename(proportion_female = homicide_count)

# calculate count by gender for family member
gender_family <- homicide_raw_long %>% 
  filter(year == 2019) %>% 
  filter(dimension == "by relationship to perpetrator") %>% 
  filter(category == "Intimate partner or family member") %>% 
  filter(indicator == "Victims of intentional homicide") %>% 
  filter(sex %in% genders) %>% 
  filter(age == "Total") %>% 
  filter(region %in% continents) %>% 
  filter(unit_of_measurement == "Counts") %>% 
  group_by(sex) %>% 
  summarize(homicide_count = sum(value))

# print proportion
percentage_all <- gender_family[1,2]/(gender_family[1,2]+gender_family[2,2]) 

percentage_all %>% 
  rename(proportion_female_family = homicide_count)

# calculate same metrics for selected countries
gender_colombia <- homicide_raw_long %>% 
  filter(year == 2019) %>%
  filter(country == "Colombia") %>% 
  filter(dimension == "Total") %>% 
  filter(indicator == "Victims of intentional homicide") %>% 
  filter(sex %in% genders) %>% 
  filter(age == "Total") %>% 
  filter(unit_of_measurement == "Counts") %>% 
  group_by(sex) %>% 
  summarize(homicide_count = sum(value))

# print proportion
percentage_colombia <- gender_colombia[1,2]/(gender_colombia[1,2]+gender_colombia[2,2]) 

percentage_colombia %>% 
  rename(proportion_female_colombia = homicide_count)


# calculate same metrics for selected countries
gender_sweden <- homicide_raw_long %>% 
  filter(year == 2019) %>%
  filter(country == "Sweden") %>% 
  filter(dimension == "Total") %>% 
  filter(indicator == "Victims of intentional homicide") %>% 
  filter(sex %in% genders) %>% 
  filter(age == "Total") %>% 
  filter(unit_of_measurement == "Counts") %>% 
  group_by(sex) %>% 
  summarize(homicide_count = sum(value))

# print proportion
percentage_sweden <- gender_sweden[1,2]/(gender_sweden[1,2]+gender_sweden[2,2]) 

percentage_sweden %>% 
  rename(proportion_female_sweden = homicide_count)

# Create variables for homicide rate country/age group
rate_colombia_49 <- 58.6 
rate_sweden_49 <- 1.53

# Calculate rate male in Colombia
rate_gian <- rate_colombia_49*2*9/10 #proportion calculation
rate_gian

# Calculate rate female in Sweden
rate_vivian <- rate_sweden_49*2*1/5 #proportion calculation
rate_vivian
```

# Prediction

## conclusion?

```{r,homicide_rate_feasibility_analysis_models&fit_graph}
# clean world bank data ------------------------
world_bank <-world_bank %>%   
  janitor::clean_names() %>% 
  mutate(value = as.double(value)) %>% 
  filter(series_name == "GDP per capita (current US$)") %>% 
  dplyr::select(code, value)

# Use read_excel to read world happiness data as dataframe
world_happiness <- read_excel(happiness.temp,
                    sheet = "Sheet1",
                    range = cell_cols("A:K")) %>% 
  janitor::clean_names()
world_happiness <- world_happiness %>% 
  filter(year == 2019)

# filter for 2019
homicide_linear <- homicide_raw %>% 
  rename(year = year_3) %>% 
  rename(homicide_rate = deaths_interpersonal_violence_sex_both_age_age_standardized_rate_4) %>% 
  rename(country = entity) %>% 
  select(country, code, year, homicide_rate) %>% 
  filter(year == 2019) %>% 
  filter(!is.na(code))

# joining tables
homicide_happiness <- inner_join(homicide_linear, world_happiness, by = c("country" = "country_name"))
homicide_happiness <- homicide_happiness %>% 
  select(-year.x, -year.y, -positive_affect, -negative_affect)

homicide_happiness_gdp <- inner_join(homicide_happiness, world_bank, by = "code")

homicide_happiness_gdp <- homicide_happiness_gdp %>% 
  rename(gdp_percap = value)

# own model
model_happiness <- lm(log1p(homicide_rate) ~ life_ladder, data = homicide_happiness_gdp)
summary(model_happiness)

reduced <- homicide_happiness_gdp %>%
  dplyr::select(-life_ladder,-country, -code)

reduced<-na.omit(reduced)

model <- lm(log1p(homicide_rate) ~ healthy_life_expectancy_at_birth + social_support + 
              freedom_to_make_life_choices + generosity+gdp_percap, data = reduced)
summary(model)

# optimizer blended out as package has same selection method
#full.model <- lm(homicide_rate ~., data = reduced)
#step.model <- stepAIC(full.model, direction = "both", 
                      #trace = FALSE)
#summary(step.model)

happiness <- homicide_happiness_gdp %>% 
  ggplot(aes(x = life_ladder, y = homicide_rate)) +
  geom_point()+
  geom_smooth(method="lm", se = F, color = "#CB102D")+
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text=element_text(color = "black", size = 12),
        text=element_text(family="Roboto"),
        plot.caption = element_text(color = "grey", hjust = 0, vjust = -5),
        )

happiness
  
ggsave(file="happiness.svg", plot=happiness, width=8, height=5)
```

# Homicide death amount in age group

## conclusion

```{r, world_homicide_cases_amount_analysis_age_group_bar_chart}
# clean & filter the data for victim and the affected age group and rename columns
victim_age_data <- victim_age_data %>%
                      janitor::clean_names() %>% 
                    # rename required columns
                      rename('above 70' = deaths_interpersonal_violence_sex_both_age_70_years_number,
                             "50 - 69" = deaths_interpersonal_violence_sex_both_age_50_69_years_number,
                             "15 - 49" = deaths_interpersonal_violence_sex_both_age_15_49_years_number,
                             "5 - 14" = deaths_interpersonal_violence_sex_both_age_5_14_years_number,
                             "under 5" = deaths_interpersonal_violence_sex_both_age_under_5_number ) %>% 
                    #filter for 2019 data
                    filter(year=="2019")

# plot to depict the cases of homicide within each age group around the world in 2019 
victim_age_data_long <- victim_age_data %>% pivot_longer(
                                cols = 4:8,
                                names_to = "age_group",
                                values_to = "count"
                              )

victim_age_data_long %>% 
  mutate(proportion = count/sum(count)) %>% 
  mutate(age_group=fct_reorder(age_group, proportion)) %>% 
  ggplot(mapping = aes(x=age_group, y=proportion, fill=age_group)) +
  geom_col()+
  theme_bw()+
  scale_fill_brewer(palette = "OrRd")

```

# Homicide death amount in age group 2

## conclusion??

```{r, world_homicide_cases_amount_analysis_age_group_ball_chart}
# Create data
data_vic <- victim_age_data_long %>% 
  group_by(age_group) %>% 
  summarise(count = sum(count))

# Add a column with the text you want to display for each bubble:
data_vic$text <- paste("name: ",data_vic$age_group, "\n", "value:", data_vic$count, "\n", "You can add a story here!")

# Generate the layout
packing <- circleProgressiveLayout(data_vic$count, sizetype='area')
data_vic <- cbind(data_vic, packing)
dat.gg <- circleLayoutVertices(packing, npoints=50)

# Make the plot with a few differences compared to the static version:
p <- ggplot() + 
  geom_polygon_interactive(data = dat.gg, aes(x,y,group = id, fill=id,tooltip = data_vic$text[id], data_id = id),colour = "black", alpha = 0.6) +
  scale_fill_viridis() +
  geom_text(data = data_vic, aes(x, y, label = gsub("Group_", "", age_group)), size=4, color="black") +
  theme_void() + 
  theme(legend.position="none", plot.margin=unit(c(0,0,0,0),"cm") ) + 
  coord_equal()+
  theme_bw()+
   theme(legend.position = "none") 

p

```

# Gender and relationship in world homicide

## Conclusion?

```{r, world_homicide_cases_amount_analysis_gender&relationship_bar_chart1}
# clean & filter the data for intentional homicide
intentional_homicide_data <- intentional_homicide_data_all %>%
                                      janitor::clean_names() %>% 
  #keeping only count data
                                      filter(unit_of_measurement=="Counts") %>% 
  #removing unknown and combined data
                                      filter(sex == c("Female","Male")) %>% 
  #keeping only overall age data 
                                      filter(age=="Total") %>% 
  #filter for 2019 data
                                      filter(year=="2019")


# plot to depict homicide share among men and women around the world in 2019
intentional_homicide_data %>% 
  mutate(proportion = value/sum(value)) %>% 
  mutate(sex=fct_reorder(sex, proportion)) %>% 
  ggplot(mapping = aes(x=sex, y=proportion, fill=sex)) +
  geom_col()+
  theme_bw()+
  scale_fill_brewer(palette = "OrRd")

```

```{r, world_homicide_cases_amount_analysis_gender&relationship_bar_chart2}
# plot to depict homicide share among men and women around the world in 2019 due to partner or family related homicides
intentional_homicide_data %>% 
  filter(category==c("Intimate partner or family member" ,"Intimate partner or family member: Family member" )) %>% 
  mutate(proportion = value/sum(value)) %>% 
  #mutate(sex=fct_reorder(sex, proportion)) %>% 
  ggplot(mapping = aes(x=sex, y=proportion, fill=sex)) +
  geom_col()+
  theme_bw()+
  scale_fill_brewer(palette = "OrRd")
```
