---
title: "Session 6: In class exercises"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
```


# Highlight one country in `gapminder


- Create a character vector of colours using `c()`. 

- Use colours `"grey90"` and `"#001e62"`. 

- Assign the vector to a name, `my_colours`

- Use the vector you just created to change the colours in the plot using `scale_color_manual()`. 

- Pass the vector of colours using the `values` argument.


```{r}



```

# `NA`s in vectors

```{r}
vector <- c(1, 2, 3, 4, NA)
sum(vector)
mean(vector)
min(vector)
max(vector)

#use `na.rm=TRUE` to make sure NAs dont screw up your work
sum(vector, na.rm=TRUE)
mean(vector, na.rm=TRUE)
min(vector, na.rm=TRUE)
max(vector, na.rm=TRUE)

```

## Writing functions
```{r}
add_one <- function(x) {
  x <- x + 1
  x
}

add_one(1)

```


## Write your own function


- Create a function called `simulated_data` that doesn't take any arguments.
- In the function body, we'll return a tibble, a dataframe with three variables: `x`, `sex`, and `age`
- For `x`, use `rnorm()` to generate 50 random numbers from a Normal distribution with a mean=0 and SD=1.
- For `sex`, use `rep()` to create 50 values of "male" and "female". 

  - Hint: You'll have to give `rep()` a character vector for the first argument. The `times` argument is how many times `rep()` should repeat the first argument, so make sure you account for that. If in doubt, type `?rep` in the console

- For `age()` use the `sample()` function to sample 50 numbers from 25 to 50 with replacement.



```{r}
set.seed = 42
sim_data <- function() {

  }
sim_data()
```


## `Passing the dots ...`

```{r}
select_gapminder <- function(...) {
  gapminder %>%
    select(...)
}
select_gapminder(pop, year)
```

- Use `...` to pass the arguments of your function, `filter_summarise()`, to filter().
- In `summarise`, get the n (count) and mean life expectancy of the filtered data set
- Check `filter_summarize()` with `year == 2002`.
- Check `filter_summarize()` with `year == 2002`, but also filter for European countries only.


```{r}
filter_summarise <- function(...) {

  
  
  
}


filter_summarise(year == 2002)
  
filter_summarise(year == 2002, continent == "Africa")
```


## `Curly curly {{}}`

```{r}
plot_hist <- function(x) {
  ggplot(gapminder, aes(x = {{x}})) + 
    geom_histogram()
}



plot_hist(lifeExp)

plot_hist(pop)

plot_hist(gdpPercap)
```


## `top_scatterplot(variable, ,year, ,n)`{.smaller}



1. Filter `gapminder` by `year` using the value of `.year` (notice the period before hand!). You do NOT need curly-curly for this, as you are passing one value for `year`. 
2. To plot the top values for whatever `variable` we pass, we have to use `fct_reorder()`. Redefine `country` so it becomes a factor ordered by `{{variable}}`.
3. Use `slice_max()`, ordering by `{{variable}}` and take the top `n` values
4. Make a scatter plot. Use `{{variable}}` for `x`. For `y`, we'll use `country` 


```{r}
top_scatterplot <- function(variable, .year, .n) {

  
}



top_scatterplot(lifeExp, 1967, 30)

top_scatterplot(pop, 1982, 10)

top_scatterplot(gdpPercap, 1997, 20)


```


## Subsetting lists

```{r}
number_list <- list(
  a = rnorm(8, mean = 2, sd = 1.4),
  b = rnorm(7, mean = 1, sd = 1),
  c = rnorm(6, mean = 10, sd = 3)
)
```


```{r}
single_bracket <- number_list["a"]
single_bracket
typeof(single_bracket)
```

```{r}
dollar <- number_list$a
dollar
typeof(dollar)
```

```{r}
double_bracket <- number_list[["a"]]
double_bracket
typeof(double_bracket)
```


## Lists and map()


```{r}
set.seed(42)
x_list <- list(x = rnorm(100), 
               y = rnorm(100), 
               z = rnorm(100))
```


```{r}
list(
  sum_x = sum(x_list$x),
  sum_y = sum(x_list$y),
  sum_z = sum(x_list$z)
)  

map(x_list, sum)
```


## Using `map()` with dataframes

```{r}
gapminder %>%
  dplyr::select_if(is.numeric) %>%
  map(sd)
```
## functions and map()


- Write a function named `mean_sd` that returns the mean and standard deviation of a numeric vector.
- Find the mean and SD of x
- Map your function to x_list

```{r}
set.seed(42)
x_list <- list(x = rnorm(100), 
               y = rnorm(100), 
               z = rnorm(100))
```

```{r}
mean_sd <- function(x) {

  }

map(x_list, mean_sd)
```


## Return types

```{r}
map(gapminder, 
    ~length(unique(.x)))
```


```{r}
map_int(gapminder, 
        ~length(unique(.x)))
```

```{r}
map_dbl(gapminder, 
        ~length(unique(.x)))
```


```{r}
map_chr(gapminder, 
        ~length(unique(.x)))
```


## `map()` and functions

```{r}
set.seed(7)
exams <- list(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)
exams
```

```{r}
exams %>% 
  map(mean)

exams %>% 
  map_dbl(mean)

```

## `set_names()` and `map_dfr()`


```{r}
names <- c("hw1","hw2","hw3","hw4","hw5","hw6","hw7","hw8","hw9","hw10") 
set_names(exams$student1, nm = names)

exams %>% 
  map(set_names, nm = names)
```

```{r}
exams %>%
  map(set_names, nm = names) %>%
  map_dfr(bind_rows)
```


## `~` and `.` to map generic expressions

- Begin the expression with `~`
- Use a `.` to indicate where inputs should go


```{r}
exams %>% 
  map_lgl(~mean(.) > 75)

exams %>% 
  map_dbl(~mean(.) > 75)



```


## Exercise

Recall the original mean of the scores
```{r}
exams %>% 
  map_dbl(mean)
```

Drop lowest score and recompute mean

(sum(`.input`) - min(`.input`)) / 9


'#
